### 2024-01-23 outputs/paper/paper.qmd
Update paper.qmd

---
title: "Exploring Respiratory Infection Trends"
subtitle: "A Data Analysis of Post-COVID-19 Impacts in Toronto (2020-2023)"
author: 
  - Hailey Jang
thanks: "Code and data are available at: https://open.toronto.ca/dataset/outbreaks-in-toronto-healthcare-institutions/."
date: January 23, 2024
date-format: long
abstract: "This research delves into the trajectory of respiratory infections in Toronto in the aftermath of COVID-19, utilizing datasets from 2020 and 2023. Simulated data on respiratory infection rates reveal a discernible surge in respiratory ailments post-pandemic. Statistical examinations are applied to compare infection rates, providing valuable insights into the changing dynamics of respiratory health. The study contributes to a holistic understanding of the enduring repercussions of COVID-19 on respiratory well-being within the Toronto population."
format: pdf
number-sections: true
bibliography: references.bib
---

### 2024-01-23 outputs/paper/paper.qmd
Update paper.qmd

@@
 ---
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+library(tidyverse)
+library(palmerpenguins)
+```
+
+# Introduction
+The aftermath of the COVID-19 pandemic prompts an in-depth examination of its enduring impact, with a particular focus on respiratory health.
+This study centers its exploration on Toronto, a diverse urban hub, utilizing datasets from pivotal years—2020 and 2023.
+Simulated data on respiratory infection rates form the basis of our analysis, uncovering discernible patterns in the prevalence of respiratory ailments post-pandemic.
+Through statistical tests, we aim to elucidate the magnitude of these changes, contributing to a nuanced understanding of the evolving landscape of respiratory health in Toronto.
+
+This research not only addresses local implications but also contributes valuable insights to the broader discourse on post-COVID-19 health dynamics.
+By unravelling the complexities of respiratory infections in Toronto, we strive to offer actionable insights for public health strategies, thereby mitigating the lasting impact of the pandemic on respiratory well-being.
+
+# Data {#sec-data}
+All the data utilized in this paper were sourced from the Toronto Open Data portal [@torontodata].
+Specifically, The dataset is titled 'Outbreaks in Toronto Healthcare Institutions'.
+The cleaning and analysis processes were executed through the use of R, a statistical programming software [@citeR], along with several libraries including `tidyverse` [@tidy], `ggplot2` [@ggplot], `dplyr` [@dp], `readr` [@read], `tibble` [@tib], `janitor` [@jan], `reshape2` [@reshape], `knitr` [@knit], `ggbeeswarm` [@gg], `ggrepel` [@repel], and `here` [@here].

### 2024-01-23 outputs/paper/paper.qmd
Update paper.qmd

--- 2024-01-23 outputs/paper/paper.qmd
@@
 ---
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+library(tidyverse)
+library(ggplot2)
+library(janitor)
+library(dplyr)
+```
+
+# Introduction
+
+The aftermath of the unprecedented COVID-19 pandemic has prompted an in-depth exploration of its enduring impact, with a particular emphasis on respiratory health.
+This study embarks on a comprehensive investigation centred in Toronto, a vibrant and diverse urban hub, utilizing datasets from pivotal years—specifically, 2020 and 2023.
+As we navigate the intricate landscape of post-pandemic health dynamics, our focus lies on understanding the nuanced trajectory of respiratory infections within this dynamic metropolitan context.
+
+Against the backdrop of the global health crisis, the city of Toronto serves as a microcosm reflecting the profound shifts in health patterns following the pandemic's peak.
+Leveraging simulated data on respiratory infection rates, we aim to unearth discernible patterns and trends that have emerged in the wake of the COVID-19 outbreak.
+Through rigorous statistical examinations, our objective is to not only quantify the changes but also to provide a detailed and nuanced portrayal of how respiratory health has evolved over this crucial period.
+
+This exploration extends beyond mere local implications, contributing valuable insights to the broader discourse on post-COVID-19 health dynamics.
+By unraveling the complexities of respiratory infections in Toronto, we strive to offer actionable insights for public health strategies.
+In doing so, we aspire to play a role in mitigating the lasting impact of the pandemic on respiratory well-being, fostering a holistic understanding that transcends geographical boundaries.
+
+## Rationale for Focus on Respiratory Infections
+To narrow our scope, we will clean the data to exclude symptoms of gastroenteric origin, focusing solely on respiratory infections.
+Subsequently, a dedicated R script will be employed to generate a graphical representation summarizing the trends in respiratory infections.
+
+# Methodology 
+## Data 
+All the data utilized in this paper were sourced from the Toronto Open Data portal [@torontodata].
+Specifically, The dataset is titled 'Outbreaks in Toronto Healthcare Institutions'.
+The cleaning and analysis processes were executed through the use of R, a statistical programming software [@citeR], along with several libraries including `tidyverse` [@tidy], `ggplot2` [@ggplot], `dplyr` [@dp], `readr` [@read], `janitor` [@jan].
+
+## Data Cleaning and Visualization 
+To isolate respiratory infections, an R script will be developed to filter out irrelevant data (Enteric, and others).
+The resultant graph will provide a clear overview of the trends in respiratory infections, laying the foundation for subsequent analyses.(@cleaned-raw-respiratory-problems).
+
+##Comparative Analysis of Causative Agents
+The causative agents under consideration for comparison between 2020 and 2023 include Rhinovirus, Metapneumovirus, Influenza A, Parainfluenza, Respiratory syncytial virus, and Enterovirus.
+Another dedicated R script will facilitate the creation of graphs, allowing for a comprehensive comparison of infection rates for each causative agent.
+
+## Respiratory Problems in Publich Health
+
+```{r}
+#| label: tbl-daily-calls
+#| tbl-cap: Sample of Respiratory Case in 2020
+#| echo: false
+
+respiratory_problems_2020 <- read.csv(here::here("outputs/data/Respiratory_Problems_Data_2020.csv"))
+
+respiratory_problems_Data_2020 <- head(respiratory_problems)
+
+knitr::kable(Sample_of_Respiratory_Case, col.names = c("Date","Type of Outbreak","Causative Agent"), align = c("l", "c", "c"))
+```
+
+#Results
+##Respiratory Infection Trends (2020 vs. 2023)
+The table 2 of respiratory infection trends post-COVID-19 reveals a noticeable increase in the overall prevalence of respiratory infections in 2023 compared to 2020.
+This sets the stage for a detailed examination of individual causative agents.
+
+##comparative Analysis of Causative Agents
+A detailed examination of individual causative agents highlights significant variations in infection rates between 2020 and 2023.
+Notably, Rhinovirus exhibits the highest increase in 2023 compared to 2020, suggesting a unique post-pandemic impact.
+
+#Discusiion 
+##Rhinovirus in 2023
+The resurgence of Rhinovirus (RV) in stark contrast to other respiratory viruses post the relaxation of social distancing measures necessitates a comprehensive analysis.
+A pivotal contributor to the heightened prevalence of RV lies in its distinctive protein capsid, conferring a heightened resistance to ethanol-containing disinfectants—a characteristic not shared by viruses with lipid envelopes.
+
+Furthermore, the efficacy of facemasks in filtering out RV is notably diminished due to its smaller size compared to influenza virus and coronavirus, potentially fueling the observed resurgence.
+The implementation of social-distancing measures, restrictions on social gatherings, and intensified hygiene protocols in response to the March 2020 COVID-19 outbreak substantially limited exposure opportunities to RV and other infectious agents.
+This led to a substantial reduction in both respiratory viral infections and healthcare utilization for asthma.
+However, the nuanced dynamics become apparent with the specific resurgence of RV infections following the fall 2020 reopening of schools.
+Unlike other respiratory viruses that remained diminished, RV exhibited a significant uptick in prevalence.
+
+The distinct impact of social measures on RV can be ascribed to its intrinsic characteristics—its small size and capsid coat may render it less susceptible to masking and disinfectants.
+The observed partial increase in asthma hospitalizations following school reopening suggests a intricate relationship between RV infections and asthma exacerbations.
+Importantly, admissions, while experiencing an uptick, remained lower than pre-COVID levels, indicating a nuanced interplay between respiratory infections, social dynamics, and asthma health outcomes.
+
+#Conclusion
+Tthe resurgence of Rhinovirus post-social distancing measures unveils a complex interplay of factors, including the unique characteristics of RV, the absence of influenza infections, and the dynamics of viral interference.
+The differential impact of social measures on various respiratory viruses underscores the need for tailored public health strategies.
+The observed partial increase in asthma hospitalizations following school reopening emphasizes the intricate relationship between respiratory infections and asthma outcomes.
+Moving forward, a nuanced understanding of these dynamics is essential for mitigating the impact of respiratory infections and optimizing public health interventions in the post-COVID era.

### 2024-01-24 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -6,25 +6,29 @@
 # License: MIT
 
 #### Workspace setup ####
-library(tidyverse)
-library(ggplot)
+library(dplyr)
 
-#### Simulate data ####
-# Set a seed for reproducibility
+# Set seed for reproducibility
 set.seed(123)
 
-# Simulate data for respiratory infections in Toronto for 2020 and 2023
-respiratory_data <- data.frame(
-  Year = rep(c(2020, 2023), each = 100),
-  Infection_Rate = c(rnorm(100, mean = 5, sd = 2), rnorm(100, mean = 7, sd = 3))
+#### Simulate data ####
+#### Simulate data ####
+# Simulate data for 2020
+data_2020 <- data.frame(
+  Type_of_Outbreak = sample(c("gastroenteric", "respiratory", "others"), 100, replace = TRUE),
+  Causative_Agent = sample(c("Rhinovirus", "Metapneumovirus", "Respiratory syncytial virus", "Influenza", "Parainfluenza", "Enterovirus"), 
+                           100, replace = TRUE)
 )
 
-#### Visualize the data with ggplot2 ####
-# Create a bar plot to compare infection rates between 2020 and 2023
-ggplot(respiratory_data, aes(x = factor(Year), y = Infection_Rate, fill = factor(Year))) +
-  geom_bar(stat = "identity", position = "dodge", color = "black") +
-  labs(title = "Comparison of Respiratory Infection Rates (2020 vs. 2023)",
-       x = "Year",
-       y = "Infection Rate") +
-  scale_fill_manual(values = c("#0072B2", "#D55E00")) +
-  theme_minimal()
+# Simulate data for 2023
+data_2023 <- data.frame(
+  Causative_Agent = sample(c("Rhinovirus", "Metapneumovirus", "Respiratory syncytial virus", "Influenza", "Parainfluenza", "Enterovirus"), 
+                           100, replace = TRUE)
+)
+
+# Save simulated data to CSV files
+write.csv(data_2020, "simulated_data_2020.csv", row.names = FALSE)
+write.csv(data_2023, "simulated_data_2023.csv", row.names = FALSE)
+
+# Print message
+cat("Simulation complete. Simulated data saved as simulated_data_2020.csv and simulated_data_2023.csv\n")

### 2024-01-24 scripts/01-download_data.R
Update 01-download_data.R

@@ -1,6 +1,6 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from OpenDataToronto
-# Author: Hailey jang 
+# Author: Hailey Jang 
 # Date: January 23, 2024
 # Contact: hailey.jang@mail.utoronto.ca
 # License: MIT
@@ -10,33 +10,33 @@ library(opendatatoronto)
 library(tidyverse)
 
 #### Download data ####
-# Got the data from opendatatoronto
-#code based from https://open.toronto.ca/dataset/outbreaks-in-toronto-healthcare-institutions/
+# Got the data from OpenDataToronto
+# Code based on https://open.toronto.ca/dataset/outbreaks-in-toronto-healthcare-institutions/
 
-#Write CSV file 
+# Read CSV file 
 # Data of respiratory problems in 2020
 raw_respiratory_problems_2020 <-
   read_csv(
     file = "https://ckan0.cf.opendata.inter.prod-toronto.ca/tl/datastore/dump/2bbbb602-f621-4f78-af5c-0d604fd63831", 
     show_col_types = FALSE
-    )
+  )
 
 # Data of respiratory problems in 2023
 raw_respiratory_problems_2023 <-
   read_csv(
     file = "https://ckan0.cf.opendata.inter.prod-toronto.ca/tl/datastore/dump/15ac28a1-ece5-4a97-8e72-11227be9f4f7",
     show_col_types = FALSE
-    )
+  )
 
 #### Save data ####
 
-# save our acquired respiratory problems rates in 2020
+# Save our acquired respiratory problems rates in 2020
 write.csv(
   x = raw_respiratory_problems_2020, 
   file = "inputs/data/Respiratory_Problems_Data_2020.csv"
 )
 
-# save our acquired respiratory problems rates in 2023
+# Save our acquired respiratory problems rates in 2023
 write.csv(
   x = raw_respiratory_problems_2023, 
   file = "inputs/data/Respiratory_Problems_Data_2023.csv"

### 2024-01-24 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -15,13 +15,14 @@ set.seed(123)
 #### Simulate data ####
 # Simulate data for 2020
 data_2020 <- data.frame(
-  Type_of_Outbreak = sample(c("gastroenteric", "respiratory", "others"), 100, replace = TRUE),
+  Type_of_Outbreak = sample(c("Enteric", "respiratory", "others"), 100, replace = TRUE),
   Causative_Agent = sample(c("Rhinovirus", "Metapneumovirus", "Respiratory syncytial virus", "Influenza", "Parainfluenza", "Enterovirus"), 
                            100, replace = TRUE)
 )
 
 # Simulate data for 2023
 data_2023 <- data.frame(
+  Type_of_Outbreak = sample(c("Enteric", "respiratory", "others"), 100, replace = TRUE),
   Causative_Agent = sample(c("Rhinovirus", "Metapneumovirus", "Respiratory syncytial virus", "Influenza", "Parainfluenza", "Enterovirus"), 
                            100, replace = TRUE)
 )

### 2024-01-24 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -12,7 +12,6 @@ library(dplyr)
 set.seed(123)
 
 #### Simulate data ####
-#### Simulate data ####
 # Simulate data for 2020
 data_2020 <- data.frame(
   Type_of_Outbreak = sample(c("Enteric", "respiratory", "others"), 100, replace = TRUE),

### 2024-01-24 scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -6,38 +6,22 @@
 # License: MIT
 
 #### Workspace setup ####
-library(tidyverse)
-library(janitor)
 library(dplyr)
 
-#### Clean data ####
+#### Read downloaded data ####
+# Assuming you have downloaded the data as CSV files
+data_2020 <- read.csv("inputs/data/Respiratory_Problems_Data_2020.csv")
+data_2023 <- read.csv("inputs/data/Respiratory_Problems_Data_2023.csv")
 
-# Read and clean the 2020 dataset
-raw_respiratory_problems_2020 <- 
-  read_csv("inputs/data/Respiratory_Problems_Data_2020.csv") %>%
-  filter(`Type of Outbreak` == "Respiratory")
+#### Data Cleaning ####
+# Filter out rows with Type_of_Outbreak 'Enteric' and 'others'
+cleaned_data_2020 <- data_2020 %>% filter(Type_of_Outbreak != "Enteric" & Type_of_Outbreak != "others")
+cleaned_data_2023 <- data_2023 %>% filter(Type_of_Outbreak != "Enteric" & Type_of_Outbreak != "others")
 
-cleaned_raw_respiratory_problems_2020 <-
-  raw_respiratory_problems_2020 %>%
-  # Add any additional data cleaning steps if needed
-  select(Enteric, other) %>%
+#### Save cleaned data ####
+# Save cleaned data to CSV files
+write.csv(cleaned_data_2020, "cleaned_data_2020.csv", row.names = FALSE)
+write.csv(cleaned_data_2023, "cleaned_data_2023.csv", row.names = FALSE)
 
-# Read and clean the 2023 dataset
-raw_respiratory_problems_2023 <- 
-  read_csv("inputs/data/Respiratory_Problems_Data_2023.csv") %>%
-  filter(`Type of Outbreak` == "Respiratory")
-
-cleaned_raw_respiratory_problems_2023 <-
-  raw_respiratory_problems_2023 %>%
-  # Add any additional data cleaning steps if needed
-  select(Enteric, other) %>%
-
-#### Save data ####
-
-# Save the cleaned 2020 dataset
-write_csv(cleaned_raw_respiratory_problems_2020, 
-          "outputs/data/cleaned_Respiratory_Problems_Data_2020.csv")
-
-# Save the cleaned 2023 dataset
-write_csv(cleaned_raw_respiratory_problems_2023, 
[…]

### 2024-01-24 scripts/03-test_data.R
Update 03-test_data.R

@@ -7,27 +7,26 @@
 
 #### Workspace setup ####
 library(tidyverse)
-library(dplyr)
 
-#### Test data ####
-# Set a seed for reproducibility
-set.seed(123)
+#### Read cleaned data ####
+cleaned_data_2020 <- read.csv("cleaned_data_2020.csv")
+cleaned_data_2023 <- read.csv("cleaned_data_2023.csv")
 
-# Simulate data for respiratory infections in Toronto for 2020 and 2023
-respiratory_data <- data.frame(
-  Year = rep(c(2020, 2023), each = 100),
-  Infection_Rate = c(rnorm(100, mean = 5, sd = 2), rnorm(100, mean = 7, sd = 3))
-)
+#### Data Cleaning Tests ####
+# Test: Rows with 'gastroenteric' should be removed
+test_that("Rows with 'gastroenteric' should be removed", {
+  expect_false(any(cleaned_data_2020$Type_of_Outbreak %in% c('gastroenteric')))
+  expect_false(any(cleaned_data_2023$Type_of_Outbreak %in% c('gastroenteric')))
+})
 
-# Save the simulated data
-write_csv(respiratory_data, "inputs/data/Respiratory_Problems_Data_Simulated.csv")
+# Test: Rows with 'others' should be removed
+test_that("Rows with 'others' should be removed", {
+  expect_false(any(cleaned_data_2020$Type_of_Outbreak %in% c('others')))
+  expect_false(any(cleaned_data_2023$Type_of_Outbreak %in% c('others')))
+})
 
-# Test the data cleaning script
-source("scripts/02-data_cleaning.R")
-
-# Print a summary or glimpse of the cleaned data
-cat("Cleaned 2020 data:\n")
-print(glimpse(cleaned_raw_respiratory_problems_2020))
-
-cat("\nCleaned 2023 data:\n")
-print(glimpse(cleaned_raw_respiratory_problems_2023))
+# Test: Number of rows in cleaned datasets should not be greater than the original datasets
+test_that("Number of rows in cleaned datasets should not be greater than the original datasets", {
+  expect_true(nrow(cleaned_data_2020) <= nrow(read.csv("inputs/data/Respiratory_Problems_Data_2020.csv")))
+  expect_true(nrow(cleaned_data_2023) <= nrow(read.csv("inputs/data/Respiratory_Problems_Data_2023.csv")))
+})

### 2024-01-25 outputs/paper/paper.qmd
Update paper.qmd

--- 2024-01-23 outputs/paper/paper.qmd
@@
 ---
-title: "Exploring Respiratory Infection Trends"
-subtitle: "A Data Analysis of Post-COVID-19 Impacts in Toronto (2020-2023)"
+title: "Unveiling Toronto's Respiratory Landscape"
+subtitle: "Analyzing Post-COVID-19 Infection Trajectory and Rhinovirus Surges through 2020 and 2023 Datasets"
 author: 
   - Hailey Jang
-thanks: "Code and data are available at: https://open.toronto.ca/dataset/outbreaks-in-toronto-healthcare-institutions/."
+thanks: "Code and data supporting this analysis is available at: https://open.toronto.ca/dataset/outbreaks-in-toronto-healthcare-institutions/."
 date: January 23, 2024
 date-format: long
-abstract: "This research delves into the trajectory of respiratory infections in Toronto in the aftermath of COVID-19, utilizing datasets from 2020 and 2023.
-Simulated data on respiratory infection rates reveal a discernible surge in respiratory ailments post-pandemic.
-Statistical examinations are applied to compare infection rates, providing valuable insights into the changing dynamics of respiratory health.
-The study contributes to a holistic understanding of the enduring repercussions of COVID-19 on respiratory well-being within the Toronto population."
+abstract: "This research investigates the impact of COVID-19 on respiratory infections in Toronto, utilizing datasets from 2020 and 2023.
+Through analysis of simulated data on respiratory infection rates, with a focus on Rhinovirus, a noticeable surge in respiratory ailments post-pandemic is revealed.
+Statistical examinations comparing infection rates offer insights into the changing dynamics of respiratory health in the city.
+This study contributes valuable information towards understanding the lasting effects of COVID-19 on respiratory well-being in the Toronto population, providing a crucial foundation for public health strategies and interventions."
 format: pdf
@@
 ---
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+library(tidyverse)
+library(ggplot2)
+library(janitor)
+library(dplyr)
+```
+
+# Introduction
+
+In the wake of the global COVID-19 pandemic, understanding the intricate repercussions on public health has become imperative.
+This research delves into the specific realm of respiratory infections in the city of Toronto, a vibrant urban center, exploring the trajectory and dynamics post the initial pandemic wave.
+The study employs datasets spanning 2020 and 2023, focusing on the prevalence of respiratory ailments, with a particular emphasis on Rhinovirus, to shed light on the enduring impacts on community health.
+
+As we navigate the aftermath of a transformative global event, there is a discernible gap in our knowledge concerning the long-term consequences of COVID-19, specifically in terms of respiratory health trends within urban populations.
+This paper aims to bridge this gap by presenting a comprehensive analysis of the simulated data on respiratory infection rates in Toronto, providing a detailed examination of the patterns and shifts observed.
+The findings illuminate a notable surge in respiratory ailments post-pandemic, with Rhinovirus emerging as a key player in this evolving landscape.
+
+The significance of this research extends beyond the empirical observations, as it offers valuable insights into the changing dynamics of respiratory health in a post-COVID-19 era.
+Unraveling these patterns not only contributes to the scientific understanding of the virus's enduring impact but also holds practical implications for public health interventions and strategic planning.
+By comprehensively addressing the evolving landscape of respiratory infections in Toronto, this paper aims to inform and guide future policies and initiatives aimed at safeguarding community well-being.
+
+To provide a roadmap for readers, this paper is structured to first elucidate the methodology employed in collecting and analyzing the data.
+Subsequently, the results section articulates the key findings, followed by a discussion that contextualizes these findings within the broader framework of public health.
+Finally, the conclusion encapsulates the paper's key takeaways, implications, and avenues for future research, offering a comprehensive overview of the study's contributions to our understanding of post-pandemic respiratory health.
+
+# Data 
+All the data used in this study were sourced form the Toronto Open Data portal [@torontodata].The dataset, titled 'Outbreaks in Toronto Healthcare Institutions' serves as the foundation for our analyses.
+
+## Respiratory Infections in 2020
+The examination of respiratory outbreak reports in Toronto healthcare institutions for the years 2020 and 2023 unveils intriguing trends that provide valuable insights into the dynamic nature of infectious diseases within these settings.
+Graph 1, representing 2020, showcases a consistent theme where respiratory infections notably outpace Enteric illnesses, reflecting the persistent challenges posed by respiratory health issues.
+The prevalence of respiratory outbreaks in this year aligns with historical expectations.
+
+```{r}
+#| echo: false
+#| include: true
+#| message: false
+#| label: Sample of Data
+#| tbl-cap: Reports of Respiratory Outbreaks (2020)
+
+library(tidyverse)
+
+# Read the data
+url <- "https://ckan0.cf.opendata.inter.prod-toronto.ca/tl/datastore/dump/2bbbb602-f621-4f78-af5c-0d604fd63831"
+respiratory_data_2020 <- read_csv(url, show_col_types = FALSE,
+                                  col_types = cols(
+                                    `Date Outbreak Began` = col_date(),
+                                    `Date Declared Over` = col_date()
+                                  ))
+
+# Select relevant columns
+selected_columns <- c("Institution Name", "Type of Outbreak", "Causative Agent - 1", "Date Outbreak Began", "Date Declared Over")
+filtered_data <- respiratory_data_2020 %>% select(all_of(selected_columns))
+
+# Plotting
+ggplot(filtered_data, aes(x = `Type of Outbreak`, fill = `Type of Outbreak`)) +
+  geom_bar() +
+  labs(title = "Reports of Respiratory Outbreaks (2020)",
+       x = "Type of Outbreak",
+       y = "Count") +
+  theme_minimal()
+
+```
+## Respiratory Infections in 2023
+Contrastingly, Graph 2, representing the post-COVID period in 2023, tells a more intricate story.
+While the baseline trend of respiratory infections surpassing Enteric illnesses persists, there is a marked escalation in the frequency of respiratory outbreaks compared to the preceding year.
+This notable surge in respiratory infection rates may be attributed to several factors, including the lingering effects of the COVID-19 pandemic.
+The classification of 2023 as post-COVID suggests a potential connection between the ongoing impact of the pandemic and the heightened prevalence of respiratory infections.
+```{r}
+#| echo: false
+#| include: true
+#| message: false
+#| label: Sample of Data 2023
+#| tbl-cap: Reports of Respiratory Outbreaks (2023)
+
+library(tidyverse)
+
+# Read the data
+url_2023 <- "https://ckan0.cf.opendata.inter.prod-toronto.ca/tl/datastore/dump/15ac28a1-ece5-4a97-8e72-11227be9f4f7"
+respiratory_data_2023 <- read_csv(url_2023, show_col_types = FALSE,
+                                  col_types = cols(
+                                    `Date Outbreak Began` = col_date(),
+                                    `Date Declared Over` = col_date()
+                                  ))
+
+# Select relevant columns
+selected_columns_2023 <- c("Institution Name", "Type of Outbreak", "Causative Agent-1", "Date Outbreak Began", "Date Declared Over")
+filtered_data_2023 <- respiratory_data_2023 %>% select(all_of(selected_columns_2023))
+
+# Plotting
+ggplot(filtered_data_2023, aes(x = `Type of Outbreak`, fill = `Type of Outbreak`)) +
+  geom_bar() +
+  labs(title = "Reports of Respiratory Outbreaks in Toronto Healthcare Institutions (2023)",
+       x = "Type of Outbreak",
+       y = "Count") +
+  theme_minimal()
+```
+
+# Results 
+## Rhinovirus
+In delving deeper into the landscape of respiratory outbreaks in Toronto healthcare institutions, a crucial aspect of our analysis involves identifying the specific viruses that contribute most significantly to these incidents.
+Understanding the predominant viral agents responsible for respiratory infections is pivotal for targeted preventive measures, effective treatment strategies, and informed public health interventions.
+In this exploration, we aim to unravel the intricate network of viral contributors, shedding light on the specific viruses that exert a substantial impact on the respiratory health dynamics within these healthcare settings.
+By discerning the key viral players, we can glean insights that will not only enhance our comprehension of the current state but also inform future strategies for mitigating the prevalence and consequences of respiratory outbreaks.
+
+The graphical representation (graph 3) comparing the rates of causative agents for respiratory outbreaks in Toronto healthcare institutions in 2020 and 2023 provides insightful observations.
+Notably, the data underscores a significant shift in the prominence of Rhinovirus as a causative agent between the two years.
+In 2023, Rhinovirus emerges as the predominant viral contributor, reporting a notably higher incidence compared to its occurrence in 2020.
+
+Specifically, in 2023, Rhinovirus stands out as the most frequently reported causative agent for respiratory outbreaks, with 77 documented cases.
+This substantial increase is particularly noteworthy when contrasted with its presence in 2020, where Rhinovirus was reported merely seven times.
+The surge in Rhinovirus cases in 2023 positions it at the forefront, signifying a noteworthy change in the landscape of respiratory infections during this period.
+
+Comparatively, in 2020, Respiratory syncytial virus took precedence as the leading causative agent, reporting 19 cases, followed closely by Influenza A with 18 cases.
+However, the dynamic shift in 2023, with Rhinovirus surpassing these previously dominant agents, suggests an evolving pattern in the prevalence of respiratory viruses.
+The observed rise in Rhinovirus cases may be indicative of changing epidemiological factors, altered patient demographics, or other environmental influences that warrant further investigation.
+
+```{r}
+#| echo: false
+#| include: true
+#| message: false
+#| label: Sample of Causative Agents
+#| tbl-cap: Rates of Causative Agents (2020 vs. 2023)
+
+library(tidyverse)
+
+# Function to read and clean data
+read_and_clean_data <- function(url, causative_agent_column) {
+  data <- read_csv(url)
+
+  # Exclude COVID-19 and Coronavirus from causative agents
+  cleaned_data <- data %>%
+    filter(!grepl("COVID-19|Coronavirus", data[[causative_agent_column]], ignore.case = TRUE))
+
+  return(cleaned_data)
+}
+
+# Read and clean data for 2020
+url_2020 <- "https://ckan0.cf.opendata.inter.prod-toronto.ca/tl/datastore/dump/2bbbb602-f621-4f78-af5c-0d604fd63831"
+cleaned_data_2020 <- read_and_clean_data(url_2020, "Causative Agent - 1")
+
+# Read and clean data for 2023
+url_2023 <- "https://ckan0.cf.opendata.inter.prod-toronto.ca/tl/datastore/dump/15ac28a1-ece5-4a97-8e72-11227be9f4f7"
+cleaned_data_2023 <- read_and_clean_data(url_2023, "Causative Agent-1")
+
+# Standardize column names
+cleaned_data_2020 <- cleaned_data_2020 %>%
+  rename(Causative_Agent = "Causative Agent - 1")
+
+cleaned_data_2023 <- cleaned_data_2023 %>%
+  rename(Causative_Agent = "Causative Agent-1")
+
+# Combine the datasets
+combined_data <- bind_rows(
+  mutate(cleaned_data_2020, Year = "2020"),
+  mutate(cleaned_data_2023, Year = "2023")
+)
+
+# Plotting
+ggplot(combined_data, aes(x = Causative_Agent, fill = Year)) +
+  geom_bar(position = "dodge") +
+  labs(title = "Rates of Causative Agents (2020 vs. 2023)",
+       x = "Causative Agent",
+       y = "Count") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+The graph highlights a substantial increase in Rhinovirus cases in 2023, positioning it as the primary causative agent for respiratory outbreaks in Toronto healthcare institutions.
+This observation prompts further exploration into the factors contributing to this shift and underscores the importance of ongoing surveillance and research to adapt public health strategies to the evolving landscape of respiratory infections.
+
+# Discussion
+## Rhinovirus in 2023
+The resurgence of Rhinovirus (RV) in 2023, particularly in comparison to other respiratory viruses following the relaxation of social distancing measures, demands a detailed examination.
+At the forefront of this resurgence is the unique protein capsid of Rhinovirus, endowing it with heightened resistance to ethanol-containing disinfectants—an attribute not shared by viruses with lipid envelopes.The efficacy of facemasks in filtering out Rhinovirus is notably compromised due to its smaller size compared to influenza virus and coronavirus, potentially contributing to the observed resurgence.
+The implementation of social-distancing measures, restrictions on social gatherings, and intensified hygiene protocols in response to the March 2020 COVID-19 outbreak significantly curtailed exposure opportunities to Rhinovirus and other infectious agents.
+This led to a substantial reduction in both respiratory viral infections and healthcare utilization for asthma.
+However, the intricacies of this dynamic become apparent with the specific resurgence of Rhinovirus infections following the fall 2020 reopening of schools.
+Unlike other respiratory viruses that remained subdued, Rhinovirus exhibited a significant uptick in prevalence.
+
+The distinctive impact of social measures on Rhinovirus can be attributed to its intrinsic characteristics—its small size and capsid coat may render it less susceptible to masking and disinfectants.
+The observed partial increase in asthma hospitalizations following school reopening suggests an intricate relationship between Rhinovirus infections and asthma exacerbations.
+Importantly, admissions, while experiencing an uptick, remained lower than pre-COVID levels, indicating a nuanced interplay between respiratory infections, social dynamics, and asthma health outcomes.
+
+# Conclusion
+The resurgence of Rhinovirus post-social distancing measures unveils a complex interplay of factors, highlighting the unique characteristics of Rhinovirus, the absence of influenza infections, and the dynamics of viral interference.
+The differential impact of social measures on various respiratory viruses underscores the necessity for tailored public health strategies.
+The observed partial increase in asthma hospitalizations following school reopening emphasizes the intricate relationship between respiratory infections and asthma outcomes.
+
+Moving forward, a nuanced understanding of these dynamics is essential for mitigating the impact of respiratory infections and optimizing public health interventions in the post-COVID era.
+This exploration not only contributes to our comprehension of Rhinovirus resurgence but also underscores the need for adaptive strategies in the ongoing battle against respiratory infections in healthcare settings.

### 2024-01-25 outputs/paper/paper.qmd
Update paper.qmd

@@
 This study contributes valuable information towards understanding the lasting effects of COVID-19 on respiratory well-being in the Toronto population, providing a crucial foundation for public health strategies and interventions."
+output: pdf_document
+number-sections: true
 format: pdf
-number-sections: true
 bibliography: references.bib
@@
 # Data 
-All the data used in this study were sourced form the Toronto Open Data portal [@torontodata].The dataset, titled 'Outbreaks in Toronto Healthcare Institutions' serves as the foundation for our analyses.
+All the data used in this study were sourced form the Toronto Open Data portal [@opendatatoronto].The dataset, titled 'Outbreaks in Toronto Healthcare Institutions' serves as the foundation for our analyses.
+
+For the purpose of data analysis, simulation, cleansing, and visualization, the statistical programming language R was judiciously employed [@citeR].
+Additionally, to facilitate these analytical processes, a suite of specialized R packages were utilized, including but not limited to here [@citeggplot2] and the comprehensive tidyverse suite [@tidyverse].
 
@@
 The examination of respiratory outbreak reports in Toronto healthcare institutions for the years 2020 and 2023 unveils intriguing trends that provide valuable insights into the dynamic nature of infectious diseases within these settings.
-Graph 1, representing 2020, showcases a consistent theme where respiratory infections notably outpace Enteric illnesses, reflecting the persistent challenges posed by respiratory health issues.
+@fig-1, representing 2020, showcases a consistent theme where respiratory infections notably outpace Enteric illnesses, reflecting the persistent challenges posed by respiratory health issues.
 The prevalence of respiratory outbreaks in this year aligns with historical expectations.
@@
 #| echo: false
-#| include: true
-#| message: false
-#| label: Sample of Data
-#| tbl-cap: Reports of Respiratory Outbreaks (2020)
+#| eval: true
+#| label: fig-1
+#| fig-cap: "Reports of Respiratory Outbreaks (2020)"
+#| warning: false
 
@@
 ```
+\newpage
 ## Respiratory Infections in 2023
-Contrastingly, Graph 2, representing the post-COVID period in 2023, tells a more intricate story.
+Contrastingly, @fig-2, representing the post-COVID period in 2023, tells a more intricate story.
 While the baseline trend of respiratory infections surpassing Enteric illnesses persists, there is a marked escalation in the frequency of respiratory outbreaks compared to the preceding year.
@@
 #| echo: false
-#| include: true
-#| message: false
-#| label: Sample of Data 2023
-#| tbl-cap: Reports of Respiratory Outbreaks (2023)
+#| eval: true
+#| label: fig-2
+#| fig-cap: "Reports of Respiratory Outbreaks (2023)"
+#| warning: false
 
@@
 ```
+##Prevalence of Respiratory Infections
+The analysis of respiratory infection trends in the aftermath of the COVID-19 pandemic reveals a noticeable increase in the overall prevalence of respiratory infections throughout the year 2023, presenting a distinct contrast with the preceding year, 2020.
+This notable rise extends beyond pre-pandemic norms, signifying a substantial transformation in the landscape of respiratory health, as highlighted in (@Maison).
+The observable surge in respiratory infections post-pandemic underscores the enduring impact of the COVID-19 pandemic on the broader respiratory health ecosystem.
+Consequently, there is a compelling need for an in-depth exploration of potential contributing factors and the development of adaptive public health strategies to effectively address the evolving challenges within the realm of respiratory infections.
 
@@
 
-The graphical representation (graph 3) comparing the rates of causative agents for respiratory outbreaks in Toronto healthcare institutions in 2020 and 2023 provides insightful observations.
+The graphical representation, @fig-3, comparing the rates of causative agents for respiratory outbreaks in Toronto healthcare institutions in 2020 and 2023 provides insightful observations.
 Notably, the data underscores a significant shift in the prominence of Rhinovirus as a causative agent between the two years.
@@
 #| echo: false
-#| include: true
-#| message: false
-#| label: Sample of Causative Agents
-#| tbl-cap: Rates of Causative Agents (2020 vs. 2023)
+#| eval: true
+#| label: fig-3
+#| fig-cap: "Rates of Causative Agents (2020 vs. 2023)"
+#| warning: false
 
@@
   geom_bar(position = "dodge") +
-  labs(title = "Rates of Causative Agents (2020 vs. 2023)",
-       x = "Causative Agent",
+  labs(x = "Causative Agent",
        y = "Count") +
@@
 
-The graph highlights a substantial increase in Rhinovirus cases in 2023, positioning it as the primary causative agent for respiratory outbreaks in Toronto healthcare institutions.
+The @fig-3 highlights a substantial increase in Rhinovirus cases in 2023, positioning it as the primary causative agent for respiratory outbreaks in Toronto healthcare institutions.
 This observation prompts further exploration into the factors contributing to this shift and underscores the importance of ongoing surveillance and research to adapt public health strategies to the evolving landscape of respiratory infections.
@@
 
-The distinctive impact of social measures on Rhinovirus can be attributed to its intrinsic characteristics—its small size and capsid coat may render it less susceptible to masking and disinfectants.
-The observed partial increase in asthma hospitalizations following school reopening suggests an intricate relationship between Rhinovirus infections and asthma exacerbations.
-Importantly, admissions, while experiencing an uptick, remained lower than pre-COVID levels, indicating a nuanced interplay between respiratory infections, social dynamics, and asthma health outcomes.
+The discernible influence of social interventions on Rhinovirus underscores the significance of its intrinsic attributes, particularly its diminutive size and capsid coat.
+According to Kreger (@Kreger), these characteristics may contribute to its reduced susceptibility to masking agents and disinfectants.
+The observed incremental surge in asthma-related hospitalizations subsequent to the reopening of schools implies a complex interconnection between Rhinovirus infections and exacerbations of asthma.
+Significantly, notwithstanding the increase in admissions, the levels remained below those observed prior to the onset of the COVID-19 pandemic, revealing a nuanced interplay among respiratory infections, societal dynamics, and health outcomes associated with asthma.
+
+\newpage
 
@@
 This exploration not only contributes to our comprehension of Rhinovirus resurgence but also underscores the need for adaptive strategies in the ongoing battle against respiratory infections in healthcare settings.
+
+# Reference

### 2024-01-25 outputs/paper/paper.qmd
Update paper.qmd

@@
   - Hailey Jang
-thanks: "Code and data supporting this analysis is available at: https://open.toronto.ca/dataset/outbreaks-in-toronto-healthcare-institutions/."
+thanks: "Code and data supporting this analysis is available at: https://github.com/Hailey-Jang/PostCOVID_Respiratory_Problems.git."
 date: January 23, 2024

### 2024-01-25 outputs/paper/paper.qmd
Update paper.qmd

@@
   - Hailey Jang
-thanks: "Code and data supporting this analysis is available at: https://github.com/Hailey-Jang/PostCOVID_Respiratory_Problems.git."
+thanks: "Code and data supporting this analysis is available at: "https://github.com/Hailey-Jang/PostCOVID_Respiratory_Problems.git"
 date: January 23, 2024

### 2024-01-25 outputs/paper/paper.qmd
Update paper.qmd

@@
   - Hailey Jang
-thanks: "Code and data supporting this analysis is available at: "https://github.com/Hailey-Jang/PostCOVID_Respiratory_Problems.git"
+thanks: "Code and data supporting this analysis is available at: https://github.com/Hailey-Jang/PostCOVID_Respiratory_Problems.git"
 date: January 23, 2024

### 2024-01-25 outputs/paper/paper.qmd
Update paper.qmd

@@
   - Hailey Jang
-thanks: "Code and data supporting this analysis is available at: https://github.com/Hailey-Jang/PostCOVID_Respiratory_Problems.git"
-date: January 23, 2024
+thanks: "Code and data supporting this analysis are available at: https://github.com/Hailey-Jang/PostCOVID_Respiratory_Problems.git"
+date: January 25, 2024
 date-format: long
@@
 library(ggplot2)
-library(janitor)
-library(dplyr)
 ```
@@
 
-In the wake of the global COVID-19 pandemic, understanding the intricate repercussions on public health has become imperative.
-This research delves into the specific realm of respiratory infections in the city of Toronto, a vibrant urban center, exploring the trajectory and dynamics post the initial pandemic wave.
-The study employs datasets spanning 2020 and 2023, focusing on the prevalence of respiratory ailments, with a particular emphasis on Rhinovirus, to shed light on the enduring impacts on community health.
+In response to the global landscape following the COVID-19 pandemic, this study thoroughly explores its impact on public health, specifically concentrating on respiratory infections in Toronto.
+Through a meticulous examination of simulated data spanning 2020 to 2023, the research provides a nuanced understanding of the trajectory post the initial pandemic wave, with a specific focus on the prevalence of Rhinovirus.
+The results highlight a significant increase in respiratory ailments, with Rhinovirus emerging as a notable contributing factor.
 
-As we navigate the aftermath of a transformative global event, there is a discernible gap in our knowledge concerning the long-term consequences of COVID-19, specifically in terms of respiratory health trends within urban populations.
-This paper aims to bridge this gap by presenting a comprehensive analysis of the simulated data on respiratory infection rates in Toronto, providing a detailed examination of the patterns and shifts observed.
-The findings illuminate a notable surge in respiratory ailments post-pandemic, with Rhinovirus emerging as a key player in this evolving landscape.
+Amidst global efforts to address the aftermath of the transformative COVID-19 pandemic, a noticeable gap exists in understanding long-term consequences, especially concerning respiratory health trends in urban populations.
+This paper aims to fill this gap by presenting a comprehensive analysis based on simulated data, scrutinizing respiratory infection rates in Toronto.
+The data section meticulously outlines data collection and analysis procedures, while the results section succinctly articulates key findings.
+The ensuing discussion contextualizes these findings within the broader framework of public health, shedding light on the evolving dynamics of respiratory health in a post-COVID-19 era.
 
-The significance of this research extends beyond the empirical observations, as it offers valuable insights into the changing dynamics of respiratory health in a post-COVID-19 era.
-Unraveling these patterns not only contributes to the scientific understanding of the virus's enduring impact but also holds practical implications for public health interventions and strategic planning.
-By comprehensively addressing the evolving landscape of respiratory infections in Toronto, this paper aims to inform and guide future policies and initiatives aimed at safeguarding community well-being.
-
-To provide a roadmap for readers, this paper is structured to first elucidate the methodology employed in collecting and analyzing the data.
-Subsequently, the results section articulates the key findings, followed by a discussion that contextualizes these findings within the broader framework of public health.
-Finally, the conclusion encapsulates the paper's key takeaways, implications, and avenues for future research, offering a comprehensive overview of the study's contributions to our understanding of post-pandemic respiratory health.
+As societies globally grapple with the intricate challenges posed by the aftermath of COVID-19, this study addresses the critical need for insights into enduring impacts on respiratory health, particularly in urban settings.
+Beyond empirical contributions, the study strives to play a proactive role in shaping future policies and basic infection control in real-life situations, serving as a guide for fortifying community health in the wake of unprecedented global events.
 
 # Data 
-All the data used in this study were sourced form the Toronto Open Data portal [@opendatatoronto].The dataset, titled 'Outbreaks in Toronto Healthcare Institutions' serves as the foundation for our analyses.
+All the data used in this study were sourced from the Toronto Open Data portal [@opendatatoronto].
+The dataset, titled 'Outbreaks in Toronto Healthcare Institutions' serves as the foundation for our analyses.
 
 For the purpose of data analysis, simulation, cleansing, and visualization, the statistical programming language R was judiciously employed [@citeR].
-Additionally, to facilitate these analytical processes, a suite of specialized R packages were utilized, including but not limited to here [@citeggplot2] and the comprehensive tidyverse suite [@tidyverse].
+Additionally, to facilitate these analytical processes, a suite of specialized R packages was utilized, including but not limited to here [@citeggplot2] and the comprehensive tidyverse suite [@tidyverse].
 
 ## Respiratory Infections in 2020
-The examination of respiratory outbreak reports in Toronto healthcare institutions for the years 2020 and 2023 unveils intriguing trends that provide valuable insights into the dynamic nature of infectious diseases within these settings.
-@fig-1, representing 2020, showcases a consistent theme where respiratory infections notably outpace Enteric illnesses, reflecting the persistent challenges posed by respiratory health issues.
-The prevalence of respiratory outbreaks in this year aligns with historical expectations.
+Examining reports on respiratory outbreaks in Toronto healthcare institutions from 2020 to 2023 unveils intriguing patterns that provide valuable insights into the dynamic nature of infectious diseases.
+In Figure 1, which represents the year 2020, a consistent trend emerges, showcasing that respiratory infections consistently surpass Enteric illnesses.
+Out of 553 reported cases, 519 are respiratory, constituting approximately 93.85% of the total cases.
+Enteric outbreaks rank as the second most common with 33 cases, making up about 5.97%.
+There is only 1 case categorized as Other, representing 0.18%, and no cases are listed under Others.
+This underscores the persistent challenges posed by respiratory health issues, with respiratory outbreaks significantly outnumbering other types.
+Considering the early onset of the COVID-19 outbreak, the prevalence of respiratory outbreaks in this year aligns with historical expectations.
 
@@
 ```
-\newpage
+
 ## Respiratory Infections in 2023
-Contrastingly, @fig-2, representing the post-COVID period in 2023, tells a more intricate story.
-While the baseline trend of respiratory infections surpassing Enteric illnesses persists, there is a marked escalation in the frequency of respiratory outbreaks compared to the preceding year.
-This notable surge in respiratory infection rates may be attributed to several factors, including the lingering effects of the COVID-19 pandemic.
-The classification of 2023 as post-COVID suggests a potential connection between the ongoing impact of the pandemic and the heightened prevalence of respiratory infections.
+In contrast, Figure 2, representing the post-COVID period in 2023, unfolds a more complex narrative.
+While the underlying trend of respiratory infections surpassing Enteric illnesses persists, there is a significant increase in the frequency of respiratory outbreaks compared to the previous year.
+In the dataset, respiratory cases total 1016, with only three cases listed under 'Others'.
+There are also 47 cases of Enteric illnesses, signifying that almost 100% of the outbreaks are of a respiratory nature.
+This notable surge in respiratory infection rates may be attributed to various factors, including the lingering effects of the COVID-19 pandemic.
+The classification of 2023 as post-COVID implies a potential connection between the ongoing impact of the pandemic and the heightened prevalence of respiratory infections.
+
 ```{r}
@@
 ```
-##Prevalence of Respiratory Infections
+## Prevalence of Respiratory Infections
 The analysis of respiratory infection trends in the aftermath of the COVID-19 pandemic reveals a noticeable increase in the overall prevalence of respiratory infections throughout the year 2023, presenting a distinct contrast with the preceding year, 2020.
 This notable rise extends beyond pre-pandemic norms, signifying a substantial transformation in the landscape of respiratory health, as highlighted in (@Maison).
-The observable surge in respiratory infections post-pandemic underscores the enduring impact of the COVID-19 pandemic on the broader respiratory health ecosystem.
 Consequently, there is a compelling need for an in-depth exploration of potential contributing factors and the development of adaptive public health strategies to effectively address the evolving challenges within the realm of respiratory infections.

### 2024-01-25 outputs/paper/paper.qmd
Update paper.qmd

@@
 In this exploration, we aim to unravel the intricate network of viral contributors, shedding light on the specific viruses that exert a substantial impact on the respiratory health dynamics within these healthcare settings.
-By discerning the key viral players, we can glean insights that will not only enhance our comprehension of the current state but also inform future strategies for mitigating the prevalence and consequences of respiratory outbreaks.
 
-The graphical representation, @fig-3, comparing the rates of causative agents for respiratory outbreaks in Toronto healthcare institutions in 2020 and 2023 provides insightful observations.
+The graphical representation, (@fig-3), comparing the rates of causative agents for respiratory outbreaks in Toronto healthcare institutions in 2020 and 2023 provides insightful observations.
 Notably, the data underscores a significant shift in the prominence of Rhinovirus as a causative agent between the two years.
 In 2023, Rhinovirus emerges as the predominant viral contributor, reporting a notably higher incidence compared to its occurrence in 2020.
-
 Specifically, in 2023, Rhinovirus stands out as the most frequently reported causative agent for respiratory outbreaks, with 77 documented cases.
 This substantial increase is particularly noteworthy when contrasted with its presence in 2020, where Rhinovirus was reported merely seven times.
-The surge in Rhinovirus cases in 2023 positions it at the forefront, signifying a noteworthy change in the landscape of respiratory infections during this period.
 
@@
 
-The @fig-3 highlights a substantial increase in Rhinovirus cases in 2023, positioning it as the primary causative agent for respiratory outbreaks in Toronto healthcare institutions.
-This observation prompts further exploration into the factors contributing to this shift and underscores the importance of ongoing surveillance and research to adapt public health strategies to the evolving landscape of respiratory infections.
-
 # Discussion
 ## Rhinovirus in 2023
-The resurgence of Rhinovirus (RV) in 2023, particularly in comparison to other respiratory viruses following the relaxation of social distancing measures, demands a detailed examination.
-At the forefront of this resurgence is the unique protein capsid of Rhinovirus, endowing it with heightened resistance to ethanol-containing disinfectants—an attribute not shared by viruses with lipid envelopes.The efficacy of facemasks in filtering out Rhinovirus is notably compromised due to its smaller size compared to influenza virus and coronavirus, potentially contributing to the observed resurgence.
-The implementation of social-distancing measures, restrictions on social gatherings, and intensified hygiene protocols in response to the March 2020 COVID-19 outbreak significantly curtailed exposure opportunities to Rhinovirus and other infectious agents.
-This led to a substantial reduction in both respiratory viral infections and healthcare utilization for asthma.
-However, the intricacies of this dynamic become apparent with the specific resurgence of Rhinovirus infections following the fall 2020 reopening of schools.
-Unlike other respiratory viruses that remained subdued, Rhinovirus exhibited a significant uptick in prevalence.
+The resurgence of Rhinovirus (RV) in 2023, especially when compared to other respiratory viruses following the relaxation of social distancing measures, requires careful scrutiny.
+At the core of this resurgence is Rhinovirus's unique protein capsid, which confers heightened resistance to ethanol-containing disinfectants—an attribute not shared by viruses with lipid envelopes.
+The efficacy of facemasks in filtering out Rhinovirus is notably compromised due to its smaller size compared to influenza virus and coronavirus, potentially contributing to the observed resurgence.
 
-The discernible influence of social interventions on Rhinovirus underscores the significance of its intrinsic attributes, particularly its diminutive size and capsid coat.
+The implementation of social-distancing measures, restrictions on social gatherings, and intensified hygiene protocols in response to the March 2020 COVID-19 outbreak significantly reduced exposure opportunities to Rhinovirus and other infectious agents, leading to a substantial reduction in both respiratory viral infections and healthcare utilization for asthma.
+However, the complexity of this dynamic becomes apparent with the specific resurgence of Rhinovirus infections following the fall 2020 reopening of schools.
+In contrast to other respiratory viruses that remained subdued, Rhinovirus exhibited a significant uptick in prevalence.
+
+The noticeable impact of social interventions on Rhinovirus underscores the significance of its intrinsic attributes, particularly its diminutive size and capsid coat.
 According to Kreger (@Kreger), these characteristics may contribute to its reduced susceptibility to masking agents and disinfectants.
-The observed incremental surge in asthma-related hospitalizations subsequent to the reopening of schools implies a complex interconnection between Rhinovirus infections and exacerbations of asthma.
-Significantly, notwithstanding the increase in admissions, the levels remained below those observed prior to the onset of the COVID-19 pandemic, revealing a nuanced interplay among respiratory infections, societal dynamics, and health outcomes associated with asthma.
+Ordinary breathing and speech generate a large amount of potentially infectious aerosols, with particles from human exhalation flows like coughing, sneezing, and breathing forming a two-phase buoyant jet emerging from the mouth and nose.
+Even after COVID-19, the use of disinfectants and sanitizers remains crucial in preventing the resurgence of Rhinovirus in our daily lives.
 
@@
 # Conclusion
-The resurgence of Rhinovirus post-social distancing measures unveils a complex interplay of factors, highlighting the unique characteristics of Rhinovirus, the absence of influenza infections, and the dynamics of viral interference.
-The differential impact of social measures on various respiratory viruses underscores the necessity for tailored public health strategies.
-The observed partial increase in asthma hospitalizations following school reopening emphasizes the intricate relationship between respiratory infections and asthma outcomes.
+The examination of Rhinovirus resurgence after social distancing measures reveals a complex interplay of factors, highlighting the unique characteristics of Rhinovirus and the intricacies of viral interference.
+The varied impact of social measures on different respiratory viruses emphasizes the need for tailored public health strategies to address specific vulnerabilities.
 
-Moving forward, a nuanced understanding of these dynamics is essential for mitigating the impact of respiratory infections and optimizing public health interventions in the post-COVID era.
-This exploration not only contributes to our comprehension of Rhinovirus resurgence but also underscores the need for adaptive strategies in the ongoing battle against respiratory infections in healthcare settings.
+Understanding these dynamics becomes crucial for mitigating the impact of respiratory infections and optimizing public health interventions in the post-COVID era.
+This exploration not only deepens our understanding of Rhinovirus resurgence but also emphasizes the urgent requirement for adaptive strategies in the ongoing battle against respiratory infections in healthcare settings.
+The insights gained from this analysis can inform the development of more targeted and resilient public health measures, fostering a comprehensive and effective approach to safeguard community health amid evolving infectious challenges.
 

### 2024-01-25 outputs/paper/paper.qmd
Update paper.qmd

@@
   geom_bar() +
-  labs(title = "Reports of Respiratory Outbreaks in Toronto Healthcare Institutions (2023)",
+  labs(title = "Reports of Respiratory Outbreaks (2023)",
        x = "Type of Outbreak",
@@
 
+\newpage
+
 # Reference

### 2024-01-25 outputs/paper/paper.qmd
Update paper.qmd

@@
 date-format: long
-abstract: "This research investigates the impact of COVID-19 on respiratory infections in Toronto, utilizing datasets from 2020 and 2023.
+abstract: "This research investigates the impact of COVID-19 on respiratory infections in Toronto, utilizing datasets in 2020 and 2023.
 Through analysis of simulated data on respiratory infection rates, with a focus on Rhinovirus, a noticeable surge in respiratory ailments post-pandemic is revealed.
-Statistical examinations comparing infection rates offer insights into the changing dynamics of respiratory health in the city.
-This study contributes valuable information towards understanding the lasting effects of COVID-19 on respiratory well-being in the Toronto population, providing a crucial foundation for public health strategies and interventions."
+Statistical examinations comparing infection rates offer insights into the changing dynamics of respiratory health.
+This study contributes towards understanding the lasting effects of COVID on respiratory well-being in the city, providing a crucial foundation for public health interventions."
 output: pdf_document
 number-sections: true
+toc: true
 format: pdf
